The Effects of the NBA COVID Bubble on the NBA Playoffs

### 2022-01-29 Manuscript/NBABubbleAnalysis.Rmd
Update NBABubbleAnalysis.Rmd
Reviewer 2 suggestion to discuss home court advantage factors in introduction

# Introduction {#sec:intro}
Home-court advantage is often discussed in sports circles as a contributing
factor to the outcome of games. It is well-known that the home team typically
benefits from some competitive edge from playing at their home-court, resulting
in a better chance of winning. Thus, the NBA playing the 2020 playoffs in a
bubble due to the COVID-19 pandemic
brought a great deal of concern for fans, teams, journalists, and others.
For example, @Aschburner discusses about anticipated effects, sharing
concerns from former players, coaches and other experts about
the potential effects of removing home-court advantage. Aschburner notes that the
NBA did make attempts to recreate the effects by putting the "home” team logo on
the court and allowing the "home" team to play crowd noise and music, but most
people doubted these small attempts would recreate a true playoff atmosphere.
During the 2020 NBA playoffs, home teams only won about 48.2% of the games. This
is lower than normal, which Aschburner claims usually floats around 60%. This
shift in the home team winning percentage surely indicates the opportunity for
thorough investigation.

>>>>

# Introduction {#sec:intro}
Home-court advantage is often discussed in sports circles as a contributing
factor to the outcome of games. It is well-known that the home team typically
benefits from some competitive edge from playing at their home-court, resulting
in a better chance of winning. Thus, the NBA playing the 2020 playoffs in a
bubble due to the COVID-19 pandemic
brought a great deal of concern for fans, teams, journalists, and others.
Among the biggest is the effect from the crowd, a home crowd has the ability
to energize home teams, help control momentum in favor of the home team and 
create a chaotic and stressful environment for away teams to play in. There 
is no way replicate these effects in the bubble conditions. Also, there are 
geographic factors that are lost like altitude and time-zone effects. There's 
evidence teams across all sports in high altitude regions can rely on away
teams coming in and struggling with lower oxygen levels at a higher altitude, 
which the home teams are accustomed to playing with \citet{Lopez}. Examples,
of high altitude teams in the NBA would be the Utah Jazz and Denver Nuggets, 
both of whom participated in the 2020 NBA bubble. Also, referees are known to
be subject to favoring home teams in large part due pressure from fans, which
is no longer in play in the bubble. Refs may still have biases, but there's
nothing to sway that bias towards the home team like normal. Any positive 
effects of the home team playing in the arena they're most comfortable in 
and living in the comfort of their own home are also lost. Travel is also 
normally discussed as a factor hurting away teams, but for the context of
the playoffs in the bubble losing this may not be significant. In the
playoffs, many times (aside from the first game of the series when the 
away team travels) both teams often travel at the same time since they're 
both going back and forth between the two cities to complete the series.

 @Aschburner discusses the anticipated effects, sharing
concerns from former players, coaches and other experts about
the potential effects of removing home-court advantage. Aschburner notes that the
NBA did make attempts to recreate the effects by putting the "home” team logo on
the court and allowing the "home" team to play crowd noise and music, but most
people doubted these small attempts would recreate a true playoff atmosphere.
During the 2020 NBA playoffs, home teams only won about 48.2% of the games. This
is lower than normal, which Aschburner claims usually floats around 60%. This
shift in the home team winning percentage surely indicates the opportunity for
thorough investigation.
